Improving the communication and transparency of stock assessment using interactive

### 2018-10-12 manuscript.Rmd
Draft paragraph on stock assessment models

The datasets used in stock assessments are constantly growing. This growth in data either stems from the continuation of long-term monitoring efforts or from the addition of new monitoring programs. As such, stock assessment biologists often have to manage large volumes of data from a variaty of sources. For instance, time series of reported landings and catch-at-age are "fishery-dependent" data that are frequently used in stock assessments. These data are often analyzed in conjunction with data from "fishery-independent" surveys that track changes in abundance and, in many cases, also monitor trends in biological factors such as age composition, growth rates, sex ratios and maturation stages. For some data-rich stocks, mark and recapture studies are also caried out to estimate movement, migration, growth rate, natural mortality, and discard mortality. All of the abovementioned datasets are multidimensional and as the volume and viriaty of these data increases, it becomes more difficult to be aware of the details of each dataset and to synthesize the results.

Synthesizing data from multiple sources presents a key challenge to stock assessment. Analyses of different data sources were traditionally carried out independently and the summaries or parameters from these analyses were used in the assessment model. This approach, however, is less than ideal because information may be lost and uncertinty may be unaccounted for when we "do statistics on the statistics" [@link1999; @maunder2013].

... Advances in software have greatly facilitated the analysis of all available data, in as raw a form as appropriate, in a single integrated analysis [@maunder2013]. ...

"We are in an age of data‐intensive science and big data, and ecologists must develop the capabilities to deal with their data." [@hampton2013]

>>>>

The data sets used in stock assessments are constantly growing. This growth in data either stems from the continuation of long-term monitoring efforts or from the addition of new monitoring programs. As such, stock assessment biologists often have to manage large volumes of data from a variety of sources. For instance, time series of reported landings and catch-at-age are "fishery-dependent" data that are frequently used in stock assessments. These data are often analyzed in conjunction with data from "fishery-independent" surveys that track changes in abundance and, in many cases, also monitor trends in biological factors such as age composition, growth rates, sex ratios and maturation stages. For some data-rich stocks, mark and recapture studies are also carried out to estimate movement, migration, growth rate, natural mortality, and discard mortality. All of the above-mentioned data sets are multidimensional and as the volume and variety of these data increases, it becomes more difficult to be aware of the details of each data set and to synthesize the results.

Synthesizing data from multiple sources presents a key challenge to stock assessment. Analyses of different data sources were traditionally carried out independently and the summaries or parameters from these analyses were used in the assessment model. This approach, however, is less than ideal because information may be lost and uncertainty may be unaccounted for when we "do statistics on the statistics" [@link1999; @maunder2013]. Such issues have largely been curtailed in contemporary stock assessments thanks to advances in software that have facilitated the analysis of all available data, in as raw a form as appropriate, in a single integrated analysis [@maunder2013]. Specifically, the application of statistical modeling tools such as JAGS [@plummer2003], AD Model Builder [@fournier2012] and Template Model Builder [@kristensen2015] allow the construction of a joint likelihood for an array of observations to, in theory, extract as much information as possible about the biological and fishery processes. However, integrated analyses are not a panacea because model misspecifications and data conflicts are an inevitable consequence of simplifying reality to a small series of equations [@maunder2017]. A potential solution to this quandary to use a superensemble model, whereby multiple models with different structures are run and their predictions are supplied as covariates to an additional superensemble model [@anderson2017]. Ensemble approaches reduce the risk of picking the "wrong" model and also expands the range of hypotheses explored [@dietterich2000]. These advances greatly improve our ability to assess the status and trends of fish populations, however, modern stock assessment biologists are now faced with the overwhelming task of understanding an ever expanding array of data inputs and model outputs.

"We are in an age of data-intensive science and big data, and ecologists must develop the capabilities to deal with their data." [@hampton2013]
